An Alarming Rate of Gendered and Racialized Police Strip Searches in the Toronto Municipal, 2020-2021

### 2024-01-23 outputs/paper/paper.qmd
refine abstract

---
title: "An Alarming Rate of Gendered and Racialized Police Strip Searches in the Toronto Municipal, 2020-2021"
author: "Quang Mai"
thanks: "Code and data are available at: https://github.com/ponolite/Police_Arrests_and_Strip_Searches/tree/main"
date: today
date-format: long
abstract: "Police enforcement, particularly arrests and strip searches, has been a topic of great contention during the COVID-19 pandemic. The question commonly raised was how certain demographics are more likely to undergo arrests and strip searches than others, and why has this skewed power dynamics been maintained for so long. Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021. This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities." WHAT WAS FOUND!!
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction

@@

Structurally, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion. First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and analysis took place (@torontoPolice). Then, all the trends and findings I discovered are shown in the Results section, while the Discussion section further assesses these findings. Ultimately, the Conclusion section summarizes the paper's main insights. In additon, it's worth noting that the data was sourced from Open Data Toronto and Toronto Police Services themselves. Thus, this dataset may already contain much racial and gender bias from the very source that this paper is questioning: the Toronto Police Services and its enforcement of strip searching (@stripSearchOpEd). Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021, yet it's still worth looking at due the alarming precedents of police power abuse.

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

@@

# Results

@@

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

@@

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

# References

>>>>

---
title: "An Alarming Rate of Gendered and Racialized Police Strip Searches in the Toronto Municipal, 2020-2021"
author: "Quang Mai"
thanks: "Code and data are available at: https://github.com/ponolite/Police_Arrests_and_Strip_Searches/tree/main"
date: today
date-format: long
abstract: "Police enforcement, particularly arrests and strip searches, has been a topic of great contention during the COVID-19 pandemic. The question commonly raised was how certain demographics are more likely to undergo arrests and strip searches than others, and why has this skewed power dynamics been maintained for so long. Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021. This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities." 
format: pdf
number-sections: true
bibliography: references.bib
---

@@

## Introduction

@@

Howevever, it's worth noting that the data was sourced from Open Data Toronto and Toronto Police Services themselves. Thus, this dataset may already contain much racial and gender bias from the very source that this paper is questioning: the Toronto Police Services and its enforcement of strip searching (@stripSearchOpEd). Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021, yet it's still worth looking at due the alarming precedents of police power abuse. Nonetheless, structurally, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion. First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and analysis took place (@torontoPolice). Then, all the trends and findings I discovered are shown in the Results section, while the Discussion section further assesses these findings. Ultimately, the Conclusion section summarizes the paper's main insights.

## Data

The data package used for analysis was sourced from the Open Data Toronto Portal under the library `opendatatoronto` [@rOpenDataToronto]. Only one dataset was retrieved to examine the racial and gender identities of Toronto police arrests from 2020 to 2021: `Police Race and Identity Based Data - Arrests and Strip Searches` [@torontoPolice]. Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra], `naniar` [@rNaniar] and `knitr` [@rKnitr]. Additional details on data extraction and cleaning processes will be specified in subsequent sections below.

```{r}
#| echo: false
#| message: false
#| label: tbl-cleaned_clinic_data
#| tbl-cap: Sample of cleaned Covid-19 clinic data

#referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
head(cleaned_clinic_data, 5) |>
  kable(
    col.names = c("ID", "Name", "Type", "Ward Name"),
    booktabs = TRUE
  )
```

@@

## Results

@@

## Discussion

## Conclusion

@@

## References
